---
title: "Bayesian Analysis"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyr)
library(tidybayes)
library(ggplot2)
library(rstan)
library(bayesplot)
library(loo)
library(patchwork)
library(GGally)
library(irr)
source('../code/utils/stan_util.R') 
source('../code/utils/TSS_util.R')
# Load the data
df_total = read.csv("../data/df_total.csv")
# Stan option
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
# Create directory to save the fitted model
resultPath= '../output/Bayesian_results'
if(!dir.exists(resultPath)){dir.create(resultPath)}
```

# simpleLO

## Posterior Distribution

Run MCMC sampling

```{r}
# Setting
modelname = 'simpleLO'
variables = c('gamma', 'alpha', 'sigmaCE')
stanname = 'DMR_simpleLogit'
Nsubj = 75
RUN = TRUE

if(RUN){
  # 1. Run MCMC sampling ===============================================
  model = stan_model(paste0('../code/stanmodel/continuous/', stanname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  # 2. Save the results ===============================================
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
  df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
  saveRDS(df_summary, paste0(resultPath, '/', modelname, '_summary.rds'))
}
```

Check MCMC convergence, estimated parameter, model fit(loo)

```{r}
# Setting
modelname = 'simpleLO'
variables = c('gamma','alpha', 'sigmaCE')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
# table
round(df_summary,3)
# parameter distribution
df_summary %>%
  ggpairs(columns= variables, title = modelname, diag = list(continuous='barDiag'))+
  theme_bw()
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
# Rhat > 1.01
cat('number of Rhat > 1.01:', sum(df_summary$max.Rhat>1.01), '\n')
print(df_summary[which(df_summary$max.Rhat>1.01),])
```

## Parameter recovery

-   gamma, alpha, sigma / r=

Make simulated data

```{r}
modelname = 'simpleLO'
variables = c('gamma', 'alpha', 'sigmaCE')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
df_sim = df_total %>% filter(Task == 2) %>% select(SubjID, P_true, x1, x2)

for(i in 1:75){
  if(i <=51){NT=330}else{NT=165}
  if(i <=51){idx_i=330*(i-1)}else{idx_i=330*51+165*(i-52)}
  # parameter setting -----------------------------------------
  gamma = df_summary[i,1]
  alpha=df_summary[i,2]
  sigma = df_summary[i,3]
  # load data of subj i----------------------------------------
  subj_i_vec = (idx_i+1):(idx_i+NT)
  p = df_sim[subj_i_vec,'P_true']
  x1 = df_sim[subj_i_vec,'x1']
  x2 = df_sim[subj_i_vec,'x2']
  # make simulated data of subj i----------------------------------------
  pip = inv_logit(gamma * logit(p)) # pi(p)
  CEpred = (pip * x1^alpha + (1-pip) * x2^alpha)^(1/alpha)
  df_sim$CE[subj_i_vec] = CEpred + rnorm(NT, 0, sigma)
}

write.csv(df_sim, file=paste0('../data/simulated/Bayesian/', modelname, '.csv'),
          row.names=FALSE)
```

Run MCMC on simulated data

```{r}
modelname = 'simpleLO'
stanname = 'DMR_simpleLogit'
variables = c('gamma', 'alpha', 'sigmaCE')
Nsubj = 75 
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/continuous/', stanname, '.stan'))
  df_sim = read.csv(paste0('../data/simulated/Bayesian/', modelname, '.csv'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti', df=df_sim)
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, paste0(resultPath, '/', modelname, '_simul.rds'))
  df_summary_rcv = MCMC_summary(fit_list, variables, convergence=FALSE)
  saveRDS(df_summary_rcv, paste0(resultPath, '/', modelname, '_simul_summary.rds'))
}
```

Visualize the parameter recovery results

```{r}
modelname = 'simpleLO'
variables = c('gamma', 'alpha', 'sigmaCE')

df_summary_rcv = readRDS(paste0(resultPath, '/', modelname, '_simul_summary.rds'))
df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))

# All subject
for(param in variables){
  x = df_summary[[param]]
  y = df_summary_rcv[[param]]
  cor = cor(x,y)
  plot(x, y, xlab='simulated parameter', ylab='recovered parameter',
       main=paste0(param, ', r = ',round(cor,2)))
  abline(a=0,b=1,col='red',lty=2)
}
```

## Test-retest reliability

Run MCMC sampling separately for session 1&2.

```{r}
modelname = 'simpleLO'
stanname = 'DMR_simpleLogit'
variables = c('gamma', 'alpha', 'sigmaCE')
Nsubj = 51
RUN = TRUE

# load data
df_s1 = df_total%>%filter(Session==1, Task==2)
df_s2 = df_total%>%filter(Session==2, Task==2)

# Run MCMC sampling
if(RUN){
  model = stan_model(paste0('../code/stanmodel/continuous/', stanname, '.stan'))
  fit_list_s1 = list(); fit_list_s2 = list()
  for(i in 1:Nsubj){
    data_list_s1 = make_stan_data(i, 'conti', df=df_s1)
    data_list_s2 = make_stan_data(i, 'conti', df=df_s2)
    fit_list_s1[[i]] = sampling(model, data_list_s1, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    fit_list_s2[[i]] = sampling(model, data_list_s2, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list_s1, paste0(resultPath, '/', modelname, '_s1.rds'))
  saveRDS(fit_list_s2, paste0(resultPath, '/', modelname, '_s2.rds'))
  df_summary_s1 = MCMC_summary(fit_list_s1, variables, convergence=FALSE)
  df_summary_s2 = MCMC_summary(fit_list_s2, variables, convergence=FALSE)
  saveRDS(df_summary_s1, paste0(resultPath, '/', modelname, '_s1_summary.rds'))
  saveRDS(df_summary_s2, paste0(resultPath, '/', modelname, '_s2_summary.rds'))
}
```

Check test-retest reliability

```{r}
modelname = 'simpleLO'
variables = c('gamma', 'alpha', 'sigmaCE')

df_summary_s1 = readRDS(paste0(resultPath, '/', modelname, '_s1_summary.rds'))
df_summary_s2 = readRDS(paste0(resultPath, '/', modelname, '_s2_summary.rds'))
# test-retest correlation
for(i in 1:length(variables)){
  x = df_summary_s1[[variables[i]]]
  y = df_summary_s2[[variables[i]]]
  df_comp = data.frame(s1= x, s2= y)
  icc_val = icc(df_comp, model = "twoway", type = "consistency", unit = "single")$value
  cor = cor(x,y)
  plot(x, y, main=paste0(variables[i],', r = ',round(cor,2), ', ICC = ',round(icc_val,2)), xlab='Session1', ylab='session2')
  abline(0,1,col='red',lty=2)
}
```

# LLO

## Posterior Distribution

Run MCMC sampling

```{r}
# Setting
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')
stanname = 'LL_DMR'
Nsubj = 75
RUN = TRUE

if(RUN){
  # 1. Run MCMC sampling ===============================================
  model = stan_model(paste0('../code/stanmodel/continuous/', stanname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  # 2. Save the results ===============================================
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
  df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
  saveRDS(df_summary, paste0(resultPath, '/', modelname, '_summary.rds'))
}
```

Check MCMC convergence, estimated parameter, model fit(loo)

```{r message=FALSE}
# Setting
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
# table
round(df_summary,3)
# parameter distribution
df_summary %>%
  ggpairs(columns= variables, title = modelname, diag = list(continuous='barDiag'))+
  theme_bw()
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
# Rhat > 1.01
cat('number of Rhat > 1.01:', sum(df_summary$max.Rhat>1.01), '\n')
print(df_summary[which(df_summary$max.Rhat>1.01),])
```

## Parameter recovery

Make simulated data

```{r}
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
df_sim = df_total %>% filter(Task == 2) %>% select(SubjID, P_true, x1, x2)

for(i in 1:75){
  if(i <=51){NT=330}else{NT=165}
  if(i <=51){idx_i=330*(i-1)}else{idx_i=330*51+165*(i-52)}
  # parameter setting -----------------------------------------
  gamma = df_summary[i,1]
  L0 = df_summary[i,2]
  alpha=df_summary[i,3]
  sigma = df_summary[i,4]
  # load data of subj i----------------------------------------
  subj_i_vec = (idx_i+1):(idx_i+NT)
  p = df_sim[subj_i_vec,'P_true']
  x1 = df_sim[subj_i_vec,'x1']
  x2 = df_sim[subj_i_vec,'x2']
  # make simulated data of subj i----------------------------------------
  pip = inv_logit(gamma * logit(p) + (1-gamma)*L0) # pi(p)
  CEpred = (pip * x1^alpha + (1-pip) * x2^alpha)^(1/alpha)
  df_sim$CE[subj_i_vec] = CEpred + rnorm(NT, 0, sigma)
}

write.csv(df_sim, file=paste0('../data/simulated/Bayesian/', modelname, '.csv'),
          row.names=FALSE)
```

Run MCMC on simulated data

```{r}
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')
stanname = 'LL_DMR'
Nsubj = 75 
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/continuous/', stanname, '.stan'))
  df_sim = read.csv(paste0('../data/simulated/Bayesian/', modelname, '.csv'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti', df=df_sim)
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, paste0(resultPath, '/', modelname, '_simul.rds'))
  df_summary_rcv = MCMC_summary(fit_list, variables, convergence=FALSE)
  saveRDS(df_summary_rcv, paste0(resultPath, '/', modelname, '_simul_summary.rds'))
}
```

Visualize the parameter recovery results

```{r}
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')

df_summary_rcv = readRDS(paste0(resultPath, '/', modelname, '_simul_summary.rds'))
df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))

# All subject
for(param in variables){
  x = df_summary[[param]]
  y = df_summary_rcv[[param]]
  cor = cor(x,y)
  plot(x, y, xlab='simulated parameter', ylab='recovered parameter',
       main=paste0(param, ', r = ',round(cor,2)))
  abline(a=0,b=1,col='red',lty=2)
}
```

## Test retest reliability

Run MCMC separately on session 1 & 2.

```{r}
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')
stanname = 'LL_DMR'
Nsubj = 51
RUN = TRUE

# load data
df_s1 = df_total%>%filter(Session==1, Task==2)
df_s2 = df_total%>%filter(Session==2, Task==2)

# Run MCMC sampling
if(RUN){
  model = stan_model(paste0('../code/stanmodel/continuous/', stanname, '.stan'))
  fit_list_s1 = list(); fit_list_s2 = list()
  for(i in 1:Nsubj){
    data_list_s1 = make_stan_data(i, 'conti', df=df_s1)
    data_list_s2 = make_stan_data(i, 'conti', df=df_s2)
    fit_list_s1[[i]] = sampling(model, data_list_s1, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    fit_list_s2[[i]] = sampling(model, data_list_s2, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list_s1, paste0(resultPath, '/', modelname, '_s1.rds'))
  saveRDS(fit_list_s2, paste0(resultPath, '/', modelname, '_s2.rds'))
  df_summary_s1 = MCMC_summary(fit_list_s1, variables, convergence=FALSE)
  df_summary_s2 = MCMC_summary(fit_list_s2, variables, convergence=FALSE)
  saveRDS(df_summary_s1, paste0(resultPath, '/', modelname, '_s1_summary.rds'))
  saveRDS(df_summary_s2, paste0(resultPath, '/', modelname, '_s2_summary.rds'))
}
```

Check the test retest reliability

```{r}
modelname = 'LLO'
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')

df_summary_s1 = readRDS(paste0(resultPath, '/', modelname, '_s1_summary.rds'))
df_summary_s2 = readRDS(paste0(resultPath, '/', modelname, '_s2_summary.rds'))
# test-retest correlation
for(i in 1:length(variables)){
  x = df_summary_s1[[variables[i]]]
  y = df_summary_s2[[variables[i]]]
  df_comp = data.frame(s1= x, s2= y)
  icc_val = icc(df_comp, model = "twoway", type = "consistency", unit = "single")$value
  cor = cor(x,y)
  plot(x, y, main=paste0(variables[i],', r = ',round(cor,2), ', ICC = ',round(icc_val,2)), xlab='Session1', ylab='session2')
  abline(0,1,col='red',lty=2)
}
```

# TSS_lambda (TSS5)

-   free parameter:

-   summed delta loo = 25012.8

-   MCMC convergence: good-moderate (Rhat\<1.01 except subj21(3.2), subj23(1.06), subj45(1.01))

## Posterior distribution

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS_lambda'
stanname = 'TSS5'
Nsubj = 75
RUN = FALSE

if(RUN){
  # 1. Run MCMC sampling ===============================================
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  # 2. Save the results ===============================================
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
  df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
  saveRDS(df_summary, paste0(resultPath, '/', modelname, '_summary.rds'))
}
```

Check MCMC convergence, estimated parameter, model fit(loo)

```{r message=FALSE}
# Setting
modelname = 'TSS_lambda'
variables = c('gamma','alpha', 'lambda')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
# table
round(df_summary,3)
# parameter distribution
df_summary %>%
  ggpairs(columns= variables, title=modelname, diag = list(continuous='barDiag'))+
  theme_bw()
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo), '\n')
# Rhat > 1.01
cat('number of Rhat > 1.01:', sum(df_summary$max.Rhat>1.01), '\n')
print(df_summary[which(df_summary$max.Rhat>1.01),])
```

## Parameter recovery

-   gamma, alpha, lambda / r= 1, 0.99, 1

Make simulated data

```{r}
modelname = 'TSS_lambda'
variables = c('gamma', 'alpha', 'lambda')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
df_sim = df_total %>% filter(Task == 2) %>% select(SubjID, P_true, x1, x2)

for(i in 1:75){
  if(i <=51){NT=330}else{NT=165}
  if(i <=51){idx_i=330*(i-1)}else{idx_i=330*51+165*(i-52)}
  # Change accordingly to the model ------------
  G_i = df_summary$gamma[i]
  A_i = df_summary$alpha[i]
  L_i = df_summary$lambda[i]
  # ---------------------------------------------
  for(j in 1:NT){
    idx = idx_i+j
    # Change accordingly to the model ---------------
    arg_list = list(gamma=G_i, alpha_gain=A_i, alpha_loss=A_i, lambda=L_i)
    df_sim$CE[idx] = TSS_simul(p= df_sim$P_true[idx], x1= df_sim$x1[idx], x2= df_sim$x2[idx],
                               V_func = V_pt, V_func_arg = arg_list, iter=1)
    # ------------------------------------------------
  }
}
write.csv(df_sim, file=paste0('../data/simulated/Bayesian/', modelname, '.csv'),
          row.names=FALSE)
```

Run MCMC sampling on a simulated data

```{r}
modelname = 'TSS_lambda'
stanname = 'TSS5'
Nsubj = 75 
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  df_sim = read.csv(paste0('../data/simulated/Bayesian/', modelname, '.csv'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS', df=df_sim)
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, paste0(resultPath, '/', modelname, '_simul.rds'))
}

```

Visualize the parameter recovery results

```{r}
modelname = 'TSS_lambda'
variables = c('gamma', 'alpha', 'lambda')

fit_list_rcv = readRDS(paste0(resultPath, '/', modelname, '_simul.rds'))
df_summary_rcv = MCMC_summary(fit_list, variables)
df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))

# All subject
for(param in variables){
  x = df_summary[[param]]
  y = df_summary_rcv[[param]]
  cor = cor(x,y)
  plot(x, y, xlab='simulated parameter', ylab='recovered parameter',
       main=paste0(param, ', r = ',round(cor,2)))
  abline(a=0,b=1,col='red',lty=2)
}
```

## Test-retest reliability

-   ICC of gamma, alpha, lambda = 0.82, 0.72, 0.84

Run MCMC sampling separately for session1 and session2.

```{r}
modelname = 'TSS_lambda'
stanname = 'TSS5'
Nsubj = 51
RUN = TRUE

# load data
df_s1 = df_total%>%filter(Session==1, Task==2)
df_s2 = df_total%>%filter(Session==2, Task==2)

# Run MCMC sampling
if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list_s1 = list(); fit_list_s2 = list()
  for(i in 1:Nsubj){
    data_list_s1 = make_stan_data(i, 'TSS', df=df_s1)
    data_list_s2 = make_stan_data(i, 'TSS', df=df_s2)
    fit_list_s1[[i]] = sampling(model, data_list_s1, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    fit_list_s2[[i]] = sampling(model, data_list_s2, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list_s1, paste0(resultPath, '/', modelname, '_s1.rds'))
  saveRDS(fit_list_s2, paste0(resultPath, '/', modelname, '_s2.rds'))
  df_summary_s1 = MCMC_summary(fit_list_s1, variables, convergence=FALSE)
  df_summary_s2 = MCMC_summary(fit_list_s2, variables, convergence=FALSE)
  saveRDS(df_summary_s1, paste0(resultPath, '/', modelname, '_s1_summary.rds'))
  saveRDS(df_summary_s2, paste0(resultPath, '/', modelname, '_s2_summary.rds'))
}
```

Check test-retest reliability

```{r}
modelname = 'TSS_lambda'
variables = c('gamma', 'alpha', 'lambda')

df_summary_s1 = readRDS(paste0(resultPath, '/', modelname, '_s1_summary.rds'))
df_summary_s2 = readRDS(paste0(resultPath, '/', modelname, '_s2_summary.rds'))
# test-retest correlation
for(i in 1:length(variables)){
  x = df_summary_s1[[variables[i]]]
  y = df_summary_s2[[variables[i]]]
  df_comp = data.frame(s1= x, s2= y)
  icc_val = icc(df_comp, model = "twoway", type = "consistency", unit = "single")$value
  cor = cor(x,y)
  plot(x, y, main=paste0(variables[i],', r = ',round(cor,2), ', ICC = ',round(icc_val,2)), xlab='Session1', ylab='session2')
  abline(0,1,col='red',lty=2)
}
```

# TSS_alpha (TSS7)

## Posterior distribution

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS_alpha'
stanname = 'TSS7'
Nsubj = 75
RUN = TRUE

if(RUN){
  # 1. Run MCMC sampling ===============================================
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  # 2. Save the results ===============================================
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
  df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
  saveRDS(df_summary, paste0(resultPath, '/', modelname, '_summary.rds'))
}
```

Check MCMC convergence, estimated parameter, model fit(loo)

```{r}
# Setting
modelname = 'TSS_alpha'
variables = c('gamma','alpha_gain', 'alpha_loss')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
# table
round(df_summary,3)
# parameter distribution
df_summary %>%
  ggpairs(columns= variables, title=modelname, diag = list(continuous='barDiag'))+
  theme_bw()
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo), '\n')
# Rhat > 1.01
cat('number of Rhat > 1.01:', sum(df_summary$max.Rhat>1.01), '\n')
print(df_summary[which(df_summary$max.Rhat>1.01),])
```

## Parameter recovery

Make simulated data

```{r message = FALSE}
modelname = 'TSS_alpha'
variables = c('gamma', 'alpha_gain', 'alpha_loss')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
df_sim = df_total %>% filter(Task == 2) %>% select(SubjID, P_true, x1, x2)

for(i in 1:75){
  if(i <=51){NT=330}else{NT=165}
  if(i <=51){idx_i=330*(i-1)}else{idx_i=330*51+165*(i-52)}
  # Change accordingly to the model ------------
  G_i = df_summary$gamma[i]
  AG_i = df_summary$alpha_gain[i]
  AL_i = df_summary$alpha_loss[i]
  # ---------------------------------------------
  for(j in 1:NT){
    idx = idx_i+j
    # Change accordingly to the model ---------------
    arg_list = list(gamma=G_i, alpha_gain=AG_i, alpha_loss=AL_i, lambda=1)
    df_sim$CE[idx] = TSS_simul(p= df_sim$P_true[idx], x1= df_sim$x1[idx], x2= df_sim$x2[idx],
                               V_func = V_pt, V_func_arg = arg_list, iter=1)
    # ------------------------------------------------
  }
}
write.csv(df_sim, file=paste0('../data/simulated/Bayesian/', modelname, '.csv'),
          row.names=FALSE)
```

Run MCMC on a simulated data

```{r}
modelname = 'TSS_alpha'
stanname = 'TSS7'
variables = c('gamma', 'alpha_gain', 'alpha_loss')
Nsubj = 75 
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  df_sim = read.csv(paste0('../data/simulated/Bayesian/', modelname, '.csv'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS', df=df_sim)
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, paste0(resultPath, '/', modelname, '_simul.rds'))
  df_summary = MCMC_summary(fit_list, variables, convergence=FALSE)
  saveRDS(df_summary, paste0(resultPath, '/', modelname, '_simul_summary.rds'))
}

```

Visualize parameter recovery result

```{r}
modelname = 'TSS_alpha'
variables = c('gamma', 'alpha_gain', 'alpha_loss')

fit_list_rcv = readRDS(paste0(resultPath, '/', modelname, '_simul.rds'))
df_summary_rcv = MCMC_summary(fit_list, variables)
df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))

# All subject
for(param in variables){
  x = df_summary[[param]]
  y = df_summary_rcv[[param]]
  cor = cor(x,y)
  plot(x, y, xlab='simulated parameter', ylab='recovered parameter',
       main=paste0(param, ', r = ',round(cor,2)))
  abline(a=0,b=1,col='red',lty=2)
}
```

## Test retest reliability

Run MCMC separately for session 1 & 2.

```{r}
modelname = 'TSS_alpha'
stanname = 'TSS7'
Nsubj = 51
RUN = TRUE

# load data
df_s1 = df_total%>%filter(Session==1, Task==2)
df_s2 = df_total%>%filter(Session==2, Task==2)

# Run MCMC sampling
if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list_s1 = list(); fit_list_s2 = list()
  for(i in 1:Nsubj){
    data_list_s1 = make_stan_data(i, 'TSS', df=df_s1)
    data_list_s2 = make_stan_data(i, 'TSS', df=df_s2)
    fit_list_s1[[i]] = sampling(model, data_list_s1, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    fit_list_s2[[i]] = sampling(model, data_list_s2, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list_s1, paste0(resultPath, '/', modelname, '_s1.rds'))
  saveRDS(fit_list_s2, paste0(resultPath, '/', modelname, '_s2.rds'))
  df_summary_s1 = MCMC_summary(fit_list_s1, variables, convergence=FALSE)
  df_summary_s2 = MCMC_summary(fit_list_s2, variables, convergence=FALSE)
  saveRDS(df_summary_s1, paste0(resultPath, '/', modelname, '_s1_summary.rds'))
  saveRDS(df_summary_s2, paste0(resultPath, '/', modelname, '_s2_summary.rds'))
}
```

Check test-retest reliability

```{r}
modelname = 'TSS_alpha'
variables = c('gamma', 'alpha_gain', 'alpha_loss')

df_summary_s1 = readRDS(paste0(resultPath, '/', modelname, '_s1_summary.rds'))
df_summary_s2 = readRDS(paste0(resultPath, '/', modelname, '_s2_summary.rds'))
# test-retest correlation
for(i in 1:length(variables)){
  x = df_summary_s1[[variables[i]]]
  y = df_summary_s2[[variables[i]]]
  df_comp = data.frame(s1= x, s2= y)
  icc_val = icc(df_comp, model = "twoway", type = "consistency", unit = "single")$value
  cor = cor(x,y)
  plot(x, y, main=paste0(variables[i],', r = ',round(cor,2), ', ICC = ',round(icc_val,2)), xlab='Session1', ylab='session2')
  abline(0,1,col='red',lty=2)
}
```

# TSS_full (TSS8)

summed delta loo = 25894.0

## Posterior distribution

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS_full'
stanname = 'TSS8'
Nsubj = 75
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Check MCMC convergence, estimated parameter, model fit(loo)

```{r message=FALSE}
# Setting
modelname = 'TSS_full'
variables = c('gamma','alpha_gain', 'alpha_loss', 'lambda')
RUN=TRUE

if(RUN){
  fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
  df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
  saveRDS(df_summary, paste0(resultPath, '/', modelname, '_summary.rds'))
}else{
  df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
}
# table
round(df_summary,3)
# parameter distribution
df_summary %>%
  ggpairs(columns= variables, title='TSS_full', diag = list(continuous='barDiag'))+
  theme_bw()
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo), '\n')
# Rhat > 1.01
cat('number of Rhat > 1.01:', sum(df_summary$max.Rhat>1.01), '\n')
print(df_summary[which(df_summary$max.Rhat>1.01),])
```

## Parameter recovery

-   gamma, alpha_gain, alpha_loss, lambda / r=

Make simulated data

```{r}
modelname = 'TSS_full'
variables = c('gamma', 'alpha_gain', 'alpha_loss', 'lambda')

df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))
df_sim = df_total %>% filter(Task == 2) %>% select(SubjID, P_true, x1, x2)

for(i in 1:75){
  if(i <=51){NT=330}else{NT=165}
  if(i <=51){idx_i=330*(i-1)}else{idx_i=330*51+165*(i-52)}
  # Change accordingly to the model ------------
  G_i = df_summary$gamma[i]
  AG_i = df_summary$alpha_gain[i]
  AL_i = df_summary$alpha_loss[i]
  L_i = df_summary$lambda[i]
  # ---------------------------------------------
  for(j in 1:NT){
    idx = idx_i+j
    # Change accordingly to the model ---------------
    arg_list = list(gamma=G_i, alpha_gain=AG_i, alpha_loss=AL_i, lambda=L_i)
    df_sim$CE[idx] = TSS_simul(p= df_sim$P_true[idx], x1= df_sim$x1[idx], x2= df_sim$x2[idx],
                               V_func = V_pt, V_func_arg = arg_list, iter=1)
    # ------------------------------------------------
  }
}
write.csv(df_sim, file=paste0('../data/simulated/Bayesian/', modelname, '.csv'),
          row.names=FALSE)
```

Run MCMC sampling on a simulated data

```{r}
modelname = 'TSS_full'
stanname = 'TSS8'
Nsubj = 75 # number of different initial values
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  df_sim = read.csv(paste0('../data/simulated/Bayesian/', modelname, '.csv'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS', df=df_sim)
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, paste0(resultPath, '/', modelname, '_simul.rds'))
}

```

Visualize the parameter recovery results

```{r}
modelname = 'TSS_full'
variables = c('gamma', 'alpha_gain', 'alpha_loss', 'lambda')

fit_list_rcv = readRDS(paste0(resultPath, '/', modelname, '_simul.rds'))
df_summary_rcv = MCMC_summary(fit_list, variables)
df_summary = readRDS(paste0(resultPath, '/', modelname, '_summary.rds'))

# All subject
for(param in variables){
  x = df_summary[[param]]
  y = df_summary_rcv[[param]]
  cor = cor(x,y)
  plot(x, y, xlab='simulated parameter', ylab='recovered parameter',
       main=paste0(param, ', r = ',round(cor,2)))
  abline(a=0,b=1,col='red',lty=2)
}
```

## Test-retest reliability

-   ICC of gamma, alpha_gain, alpha_loss, lambda = 0.83, 0.72, 0.69, 0.85

Compute MLE separately for session1 and session2.

```{r}
modelname = 'TSS_full'
stanname = 'TSS8'
Nsubj = 51
RUN = TRUE

# load data
df_s1 = df_total%>%filter(Session==1, Task==2)
df_s2 = df_total%>%filter(Session==2, Task==2)

# Run MCMC sampling
if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list_s1 = list(); fit_list_s2 = list()
  for(i in 1:Nsubj){
    data_list_s1 = make_stan_data(i, 'TSS', df=df_s1)
    data_list_s2 = make_stan_data(i, 'TSS', df=df_s2)
    fit_list_s1[[i]] = sampling(model, data_list_s1, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    fit_list_s2[[i]] = sampling(model, data_list_s2, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list_s1, paste0(resultPath, '/', modelname, '_s1.rds'))
  saveRDS(fit_list_s2, paste0(resultPath, '/', modelname, '_s2.rds'))
}
```

Check test-retest reliability

```{r}
modelname = 'TSS_full'
variables = c('gamma', 'alpha_gain','alpha_loss', 'lambda')

fit_list_s1 = readRDS(paste0(resultPath, '/', modelname, '_s1.rds'))
fit_list_s2 = readRDS(paste0(resultPath, '/', modelname, '_s2.rds'))
df_summary_s1 = MCMC_summary(fit_list_s1, variables)
df_summary_s2 = MCMC_summary(fit_list_s2, variables)

# test-retest correlation
for(i in 1:length(variables)){
  x = df_summary_s1[[variables[i]]]
  y = df_summary_s2[[variables[i]]]
  df_comp = data.frame(s1= x, s2= y)
  icc_val = icc(df_comp, model = "twoway", type = "consistency", unit = "single")$value
  cor = cor(x,y)
  plot(x, y, main=paste0(variables[i],', r = ',round(cor,2), ', ICC = ',round(icc_val,2)), xlab='Session1', ylab='session2')
  abline(0,1,col='red',lty=2)
}


```

## Within block reliability

```{r}
modelname = 'TSS_full'
stanname = 'TSS8'
Nsubj = 5
RUN = TRUE

# load data
df_b1 = df_total%>%filter(Block==1, Task==2)
df_b2 = df_total%>%filter(Block==2, Task==2)

# Run MCMC sampling
if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', stanname, '.stan'))
  fit_list_b1 = list(); fit_list_b2 = list()
  for(i in 1:Nsubj){
    data_list_b1 = make_stan_data(i, 'TSS', df=df_b1)
    data_list_b2 = make_stan_data(i, 'TSS', df=df_b2)
    fit_list_b1[[i]] = sampling(model, data_list_b1, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    fit_list_b2[[i]] = sampling(model, data_list_b2, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list_b1, paste0(resultPath, '/', modelname, '_b1.rds'))
  saveRDS(fit_list_b2, paste0(resultPath, '/', modelname, '_b2.rds'))
}
```

```{r}
modelname = 'TSS_full'
variables = c('gamma', 'alpha_gain','alpha_loss', 'lambda')

fit_list_b1 = readRDS(paste0(resultPath, '/', modelname, '_b1.rds'))
fit_list_b2 = readRDS(paste0(resultPath, '/', modelname, '_b2.rds'))
df_summary_b1 = MCMC_summary(fit_list_b1, variables)
df_summary_b2 = MCMC_summary(fit_list_b2, variables)

# test-retest correlation
for(i in 1:length(variables)){
  x = df_summary_b1[[variables[i]]]
  y = df_summary_b2[[variables[i]]]
  df_comp = data.frame(b1= x, b2= y)
  icc_val = icc(df_comp, model = "twoway", type = "consistency", unit = "single")$value
  cor = cor(x,y)
  plot(x, y, main=paste0(variables[i],', r = ',round(cor,2), ', ICC = ',round(icc_val,2)), xlab='Session1', ylab='session2')
  abline(0,1,col='red',lty=2)
}


```
