---
title: "Continuous models"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyr)
library(tidybayes)
library(ggplot2)
library(rstan)
library(bayesplot)
library(loo)
library(patchwork)
# Load the data
df_total = read.csv("../data/df_total.csv")
source("../code/utils/stan_util.R")
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
resultPath= '../output/continuous_modelfit'
if(!dir.exists(resultPath)){dir.create(resultPath)}
```

# 1. Expected Utility

$$
\begin{align*}
\text{CE_pred} &= U^{-1}\left( pU(x_1) + (1-p)U(x_2)\right) + \epsilon_\text{CE}\\
U(x) &= x^{\alpha}
\end{align*}
$$

-   free parameter: $\alpha, \sigma_\text{CE}$

-   Prior: alpha, sigma \~ cauchy(0,5)

-   Summed loo(S01-S05) = -8561

-   MCMC convergence = good (Rhat\<1.01)

Run MCMC sampling

```{r}
# Setting
modelname = 'EU_DMR'
Nsubj = 5
RUN = FALSE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'EU_DMR'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('alpha', 'sigmaCE')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 2. Linear Logit

$$
\begin{align*}
\pi(p) &= \text{inv_logit}(\gamma\cdot\text{logit}(p) + (1-\gamma)\cdot L0)\\
\text{CE_pred} &= (\pi(p)x_1^\alpha + (1-\pi(p))x_2^\alpha)^{1/\alpha} + \epsilon_\text{CE}
\end{align*}
$$

-   free parameter: gamma, L0, alpha, sigma

-   Prior: L0 \~ N(0,1) / alpha, gamma \~ exp(N(0,1)) / sigmaCE \~ cauchy(0,5)

-   Summed loo(S01-S05) = -7762

-   MCMC convergence = good (Rhat\<1.01)

-   **high correlation between L0 and alpha**

Run MCMC sampling

```{r}
# Setting
modelname = 'LL_DMR'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'LL_DMR'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma', 'L0', 'alpha', 'sigmaCE')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 3. LL no alpha

$$
\begin{align*}
\pi(p) &= \text{inv_logit}(a\cdot\text{logit}(p) + b)\\
\text{CE_pred} &= \pi(p)x_1 + (1-\pi(p))x_2 + \epsilon_\text{CE}\\
\end{align*}
$$

-   free parameter: a, b, sigmaCE

-   Prior: a, b \~ normal(0,1) / sigmaCE \~ cauchy(0,5)

-   Summed loo(S01-S05) = -7782

-   MCMC convergence = good (Rhat\<1.01)

-   moderate correlation between a and b

Run MCMC sampling

```{r}
# Setting
modelname = 'LLnoal_DMR'
Nsubj = 5
RUN = FALSE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'LLnoal_DMR'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('a', 'b', "sigmaCE")

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 4. BLO

$$
\begin{align*}
\lambda(p) &= \text{clip }\text{logit}(p)\text{ between }[\Delta^{-}, \Delta{+}]\ \Rightarrow \text{linear transformation between }[-\tau\Delta,+ \tau\Delta]\quad&&(\text{Thurstone scale})\\
w_p&=\frac{1}{1+\kappa p(1-p)}\quad &&(\text{variance compensation})\\
\pi(p) &= \text{inv_logit}(w_p\cdot\lambda(p)+(1-w_p)\cdot L0)
\end{align*}
$$

-   free parameter: tau, center, delta, kappa, L0, alpha, sigma

-   Prior

    -   tau, kappa, alpha \~ exp(normal(0,1)) / sigma \~ cauchy(0,5)

    -   L0 \~ (2\*inv_logit(normal(0,1))-1) \* 10

    -   center \~ (2\*inv_logit(normal(0,1))-1) \* 4.59

    -   delta \~ inv_logit(normal(0,1)) \* 4.59

-   Summed loo(S01-S05) = -7753

-   **MCMC convergence = poor (Rhat \< 1.14)**

-   **Strong correlation across parameters, multimodal distribution**

Run MCMC sampling

```{r}
# Setting
modelname = 'BLO_DMR_reparam'
Nsubj = 5
RUN = FALSE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'BLO_DMR_reparam'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c("center", "Delta", "tau", "kappa", "L0")

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 5. BLO softclip

$$
\begin{align*}
\lambda(p) &= \tau\Delta\cdot\tanh\left(\frac{\text{logit}(p)-center}{\Delta}\right)\quad&&(\text{Thurstone scale})\\
w_p&=\frac{1}{1+\kappa p(1-p)}\quad &&(\text{variance compensation})\\
\pi(p) &= \text{inv_logit}\left(w_p\cdot\lambda(p)+(1-w_p)\cdot center\right)
\end{align*}
$$

-   free parameter: tau, center, delta, kappa, L0, alpha, sigma

-   Prior = same as BLO model

-   Summed loo(S01-S05) = -7756

-   **MCMC convergence = poor (Rhat \< 3.8)**

-   **Strong correlation across parameters, multimodal distribution**

Run MCMC sampling

```{r}
# Setting
modelname = 'BLO_DMR_softclip_reparam'
Nsubj = 5
RUN = FALSE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'BLO_DMR_softclip_reparam'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c("center", "Delta", "tau", "kappa", "L0")

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 6. BLO no kappa

Softclip + no variation compensation(no kappa)

-   free parameter: tau, Deltaminus, Delta, L0, alpha, sigma

-   Prior:

    -   DeltaMinus \~ normal(0, 4.59)

    -   Delta \~ normal(2.295, 2.295)

    -   L0 \~ normal(0,10)

    -   tau, alpha \~ exp(normal(0,1)

    -   sigma \~ cauchy(0,5)

-   Summed loo(S01-S05) = -7824

-   MCMC convergence = good-moderate (Rhat \< 1.02)

-   **Strong correlation across parameters**

Run MCMC sampling

```{r}
# Setting
modelname = 'BLO_DMR_nok'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'BLO_DMR_nok'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c("DeltaMinus", "DeltaPlus", "tau_raw", "L0", "sigmaCE", "alpha_raw")

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 7. BLO no kappa - revised

$$
\begin{align*}
\pi(p) &= \text{inv_logit}\left(\tau\Delta\cdot\tanh\left(\frac{\text{logit}(p)-center}{\Delta}\right)+center\right)
\end{align*}
$$

-   BLO no kappa model with few modifications

    -   1\. "L0" is fixed to "center" so one free parameter is reduced

    -   2\. Use pi(1-p) instead of (1-pi(p)) when calculating predicted CE

<!-- -->

-   Summed loo(S01-S05) = -7783

-   MCMC convergence = good (Rhat \< 1.01)

-   moderate correlation between variables

Run MCMC sampling

```{r}
# Setting
modelname = 'BLO_DMR_nok_re'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'BLO_DMR_nok_re'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c('center', 'Delta', 'tau_raw', 'alpha_raw', 'sigmaCE')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 8. BLO no clipping

No clipping(No Delta). Scaling + variance compensation with reparameterization.

However, this model ignores the case when gamma \> 1.

$$
\begin{align*}
\pi(p) &= \frac{\text{logit}(p) -L0}{A'+B'\cdot p\cdot (1-p)}+L0\quad\text{where}\ A',B'\geq 1\\
or\quad &=\frac{\text{logit}(p)-L0}{\exp(\exp(A))+\exp(\exp(B))\cdot p\cdot (1-p)}+L0
\end{align*}
$$

-   free parameter: A, B, L0, alpha, sigma

-   Prior: A, B \~ N(0,1) / L0 \~ N(0,10) / alpha \~ exp(N(0,1)) / sigma \~ cauchy(0,5)

-   Summed loo(S01-S05) = -7764

-   MCMC convergence = good (Rhat \< 1.01)

-   moderate correlation between parameters

Run MCMC sampling

```{r}
# Setting
modelname = 'BLO_DMR_nod'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'BLO_DMR_nod'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c('A', 'B', 'alpha_raw', 'L0', 'sigmaCE')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 9. LL - revised

Linear Logit model with a modification: use pi(1-p) instead of (1-pi(p)) when calculating predicted CE

-   free parameter: a, b, alpha, sigma

<!-- -->

-   Prior: a, b \~ N(0,1) / alpha \~ exp(N(0,1)) / sigma \~ cauchy(0,5)

-   Summed loo(S01-S05) = -7770

-   MCMC convergence = good (R\<1.01)

-   **Correlation between b and alpha has reduced but there the correlation still remains moderately**

Run MCMC sampling

```{r}
# Setting
modelname = 'LL_DMR_re'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'LL_DMR_re'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c('a', 'b', 'alpha_raw', 'sigmaCE')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# 10. simple LL

$$
\begin{align*}
\pi(p) &= \text{inv_logit}(\gamma \cdot \text{logit}(p))\\
CE  &= (\pi(p) \cdot x_1^\alpha + (1-\pi(p))\cdot x_2^\alpha)^{1/\alpha} + \sigma_{CE}
\end{align*}
$$

-   free parameter: gamma, alpha, sigma

-   prior: gamma, alpha \~ exp(N(0,1)) / sigma \~ cauchy(0,5)

-   Summed loo(S01-S05) = -7780

-   MCMC convergence = good (R\<1.01)

-   mild correlation between alpha, gamma (only at S05)

Run MCMC

```{r}
# Setting
modelname = 'DMR_simpleLogit'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(file = paste0('../code/stanmodel/continuous/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'conti')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'DMR_simpleLogit'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables =  c('gamma' ,'alpha', 'sigmaCE')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```
