---
title: "Two Stage Selection model"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyr)
library(tidybayes)
library(ggplot2)
library(rstan)
library(bayesplot)
library(loo)
library(patchwork)
# Load the data
df_total = read.csv("../data/df_total.csv")
source("../code/utils/stan_util.R")
source("../code/utils/TSS_util.R")
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
resultPath= '../output/TSS_modelfit'
if(!dir.exists(resultPath)){dir.create(resultPath)}
```

# TSS model

## What is TSS model

The value function is defined based on the prospect theory

$$
\begin{align*}
V(Y;x_1, x_2, p) &= \text{value of: win } (x_1-Y) \text{ with prob }p\quad \text{or}\quad \text{loose } (Y-x_2) \text{with prob }1-p
\end{align*}
$$

**Step 1: Coarse selection**

The model choose $CE_c$ between five candidates(Y_c) between $\left[x_2, x_1\right]$$$Y_c = x_2 + (x_1-x_2)\cdot y_c\quad\text{where}\quad y_c \in \{0.1, 0.3, 0.5, 0.7, 0.9\},\quad $$

The model stochastically choose Y_c with smallest $|V(Y_c)|$ via softmax selection.

**Step 2: Fine selection**

Similar choice is made between $[CE_c-0.04\cdot2\cdot(x_1-x_2), CE_c+0.04*2*(x_1-x_2)]$

The chosen candidate become a CE.

## Value function

Every TSS model except(TSS6) uses the following value function

$$
V(Y; x_1, x_2, p) = (x_1-Y)^{\alpha_{\text{gain}}} - \lambda(Y-x_2)^{\alpha_{\text{loss}}}
$$

-   The model uses a parameter alpha assumes alpha_gain = alpha_loss = alpha

-   The model without parameter lambda assumes lambda=1

-   The parameter beta refers to reverse temperature of the softmax function.

In the continuous model we modeled the CE as follows

$$
CE = (\pi(p)\cdot x_1^\alpha + (1-\pi(p))\cdot x_2^{\alpha})^{1/\alpha} + \epsilon
$$

TSS model's usage of prospect theory fundamentally differs from the continuous model.

\<Note\> Since it is the discrete model, we can choose random selection model as the baseline model. The loo value of the baseline model is -1000. '**Summed delta loo**' refers to the total sum of (model loo - baseline loo).

# TSS1 (B)

-   free param = beta

-   Prior: beta \~ exp(N(0,1))

-   summed delta loo = -78 (worse than baseline)

-   MCMC convergence = good (Rhat\<1.01)

Simulation

```{r}
beta=0.01
arg_list = list(V_func=V_pt, beta = beta, iter = 100)
plot_list = plot_simul(TSS_simul, arg_list)
p = wrap_plots(plot_list, ncol=4)+
    plot_annotation(title=paste0('TSS1'), subtitle=paste0('beta = ',beta))
print(p)
```

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS1'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'TSS1'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('beta')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
```

# TSS2 (G,B)

-   free param = gamma, beta

-   Prior: gamma, beta \~ exp(N(0,1))

-   summed delta loo = -35 (worse than baseline)

-   MCMC convergence = good (Rhat \< 1.01)

Simulation

```{r}
# simulated data vs. observed data of S02
beta = exp(-3.44)
gamma = exp(-0.0081)
arg_list = list(V_func=V_pt, V_func_arg = list(gamma=gamma), beta = beta, iter = 100)
plot_list = plot_simul(TSS_simul, arg_list, obs_point_ID = 2)
p = wrap_plots(plot_list, ncol=4)+
    plot_annotation(title=paste0('TSS2'), subtitle=paste0('beta = ',round(beta,3), ' gamma = ', round(gamma,3)))
print(p)
```

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS2'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'TSS2'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma', 'beta')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# TSS3 (G, A, B)

-   free parameter = gamma, alpha, beta

-   prior: gamma, alpha, beta \~ exp(N(0,1))

-   summed delta loo = 403

-   MCMC convergence = poor (Rhat \> 1.7 at subj 3)

-   **Strong correlation between beta and alpha**

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS3'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'TSS3'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma','alpha', 'beta')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# TSS4 (G, A)

-   free parameter = gamma, alpha

-   prior: gamma, alpha \~ exp(N(0,1))

-   summed delta loo = 304

-   MCMC convergence = poor (Rhat \> 2 at subj 3)

```{r}
# Setting
modelname = 'TSS4'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'TSS4'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma','alpha')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# TSS5 (G, A, L)

-   free parameter = gamma, alpha, lambda

-   prior: gamma, alpha, lambda \~ exp(N(0,1))

-   summed delta loo = 966

-   MCMC convergence = good (R\<1.01)

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS5'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'TSS5'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma','alpha', 'lambda')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

Simulation

```{r}
arg_list = list(V_func=V_pt,
                V_func_arg = list(gamma=0.02983, alpha_gain=0.4663, lambda=0.889), iter = 100)
plot_list = plot_simul(TSS_simul, arg_list, obs_point_ID = 73)
p = wrap_plots(plot_list, ncol=4)+
    plot_annotation(title=paste0('TSS5'), subtitle='subj73')
print(p)

arg_list = list(V_func=V_pt,
                V_func_arg = list(gamma=242, alpha_gain=2.387e-18, lambda=77545), iter = 100)
plot_list = plot_simul(TSS_simul, arg_list, obs_point_ID = 73)
p = wrap_plots(plot_list, ncol=4)+
    plot_annotation(title=paste0('TSS5'), subtitle='subj73-recovered')
print(p)
```

# TSS6

This model uses the concept used in the continuous models. In continuous models we calculated modeled CE as follows

$$
\text{CE}_{\text{pred}}(p, x_1, x_2 ; \gamma, \alpha) = (\pi(p)\cdot x_1^\alpha + (1-\pi(p))\cdot x_2^{\alpha})^{1/\alpha}
$$

The value function of this model is

$$
V(Y;p, x_1, x_2) = |Y - \text{CE}_{\text{pred}}(p, x_1, x_2 ; \gamma, \alpha)|
$$

-   free parameter = gamma, alpha, beta

-   prior: gamma, alpha, beta \~ exp(N(0,1))

-   summed delta LOO = 463

-   MCMC convergence = good (Rhat\<1.01)

-   moderate correlation between alpha and gamma

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS6'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, convergence diagnosis, generalizability

```{r}
# Setting
modelname = 'TSS6'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma','alpha', 'beta')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=variables, off_diag_args = list(size=1, alpha=0.1)))
}
```

# TSS7 (G, AG, AL)

-   free parameter = gamma, alpha_gain, alpha_loss

    -   alpha_gain = exp(alpha_raw[1])

    -   alpha_loss = exp(alpha_raw[1]+alpha_raw[2])

-   prior: gamma \~ exp(N(0,1)) / alpha_raw \~ N(0,1)

-   summed delta LOO = 1026

-   MCMC convergence = good (Rhat \< 1.01)

-   **Moderate correlation between alpha_raw[1] and alpha_raw[2] (at subj3)**

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS7'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, MCMC convergence, loo

```{r}
# Setting
modelname = 'TSS7'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma','alpha_gain', 'alpha_loss')
vars_reparam = c('gamma', 'alpha_raw[1]', 'alpha_raw[2]')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=vars_reparam, off_diag_args = list(size=1, alpha=0.1)))
}
```

# TSS8 (G, AG, AL, L)

-   Distance = abs

-   free parameter = gamma, alpha_gain, alpha_lose, lambda

-   prior: lambda \~ exp(N(0,1)) / gamma, alpha_gain/loss -\> same as TSS7

-   summed delta loo = 1067

-   MCMC convergence = good (Rhat \< 1.01)

-   **Strong correlation between alpha_raw[2] and lambda**

Run MCMC sampling

```{r}
# Setting
modelname = 'TSS8'
Nsubj = 5
RUN = TRUE

if(RUN){
  model = stan_model(paste0('../code/stanmodel/TSS/', modelname, '.stan'))
  fit_list = list()
  for(i in 1:Nsubj){
    data_list = make_stan_data(i, 'TSS')
    fit_list[[i]] = sampling(model, data_list, chains = 4,iter = 2000, warmup = 1000,
                             seed = 1234, refresh=0)
    message(paste("✓ Subj", i, "completed."))
  }
  saveRDS(fit_list, file = paste0(resultPath,'/', modelname, '.rds'))
}
```

Posterior mean/distribution, MCMC convergence, loo

```{r}
# Setting
modelname = 'TSS8'
fit_list = readRDS(paste0(resultPath,'/', modelname, '.rds'))
variables = c('gamma','alpha_gain', 'alpha_loss', 'lambda')
vars_reparam = c('gamma', 'alpha_raw[1]', 'alpha_raw[2]', 'lambda')

# Posterior mean & convergence diagnosis
df_summary = MCMC_summary(fit_list, variables, modelcomp=TRUE)
df_summary
# Summed loo
cat('summed loo:', sum(df_summary$loo), '\n')
cat('summed delta loo:', sum(df_summary$loo) + 1000 * length(df_summary$loo))
# Trace plot
for(i in 1:length(fit_list)){
  print(traceplot(fit_list[[i]], pars=c(variables, 'lp__')))
}
# Posterior distribution
for(i in 1:5){
  fit = fit_list[[i]]
  print(mcmc_pairs(fit, pars=vars_reparam, off_diag_args = list(size=1, alpha=0.1)))
}
```

5 rows
